Informing forest carbon inventories under the Paris Agreement using the Global Forest Carbon Database (ForC v4.0)

### 2023-05-23 doc/manuscript/ESSDmanuscript.Rmd
exclude R_het_soil, knit in new display items
closes #51
closes #45

n_ind_records_EFDB_variables= C_variables[length(C_variables[,"variable"]), "n independent records in ForC"] #n independent records relevant to EFDB
n_ind_records_agb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag"]
n_ind_records_agb_components <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_foliage"]
n_ind_records_biomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass"]
n_ind_records_rootbiomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_coarse"]
n_ind_records_deadwood <- C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_down"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_standing"]
n_ind_records_OL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "litter"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "O.horizon"]
n_ind_records_TotEcosystemC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "total.ecosystem_2"] 
n_ind_records_SOC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "SOM / SOC"] 
n_ind_records_deltaagb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.agb"] 
n_ind_records_deltaOL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.O.horizon"] 
flux_variables_with_records <- c("NPP_woody", "ANPP_woody", "ANPP_woody_stem", "ANPP_woody_branch", "woody.mortality_ag", "BNPP_root", "BNPP_root_fine","BNPP_root.turnover_fine","BNPP_root_coarse" , "ANPP_litterfall", "NPP_litter", "R_het_litter", "R_het_soil")
n_ind_records_flux <- sum(C_variables$`n independent records in ForC`[C_variables$`variable` %in% flux_variables_with_records])

@@

```{r fig_variable_mapping, out.width = "14cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon."}
#this is for a two-column figure. for one-column figure, use out.width = "8.3cm"
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

>>>>

n_ind_records_EFDB_variables= C_variables[length(C_variables[,"variable"]), "n independent records in ForC"] #n independent records relevant to EFDB
n_ind_records_agb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag"]
n_ind_records_agb_components <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_woody"] + C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_ag_foliage"]
n_ind_records_biomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass"]
n_ind_records_rootbiomass <- C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_fine"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "biomass_root_coarse"]
n_ind_records_deadwood <- C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_down"]+C_variables$`n independent records in ForC`[C_variables$`variable` == "deadwood_standing"]
n_ind_records_OL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "litter"] +C_variables$`n independent records in ForC`[C_variables$`variable` == "O.horizon"]
n_ind_records_TotEcosystemC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "total.ecosystem_2"] 
n_ind_records_SOC <- C_variables$`n independent records in ForC`[C_variables$`variable` == "SOM / SOC"] 
n_ind_records_deltaagb <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.agb"] 
n_ind_records_deltaOL <- C_variables$`n independent records in ForC`[C_variables$`variable` == "delta.O.horizon"] 
flux_variables_with_records <- c("NPP_woody", "ANPP_woody", "ANPP_woody_stem", "ANPP_woody_branch", "woody.mortality_ag", "BNPP_root", "BNPP_root_fine","BNPP_root.turnover_fine","BNPP_root_coarse" , "ANPP_litterfall", "NPP_litter", "R_het_litter")
n_ind_records_flux <- sum(C_variables$`n independent records in ForC`[C_variables$`variable` %in% flux_variables_with_records])

@@

```{r fig_variable_mapping, out.width = "14cm", echo = FALSE, fig.cap = "\\textbf{Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon. **This variable is techically EFDB-relevant but not selected for submission because their is no corresponding influx variable."}
#this is for a two-column figure. for one-column figure, use out.width = "8.3cm"
knitr::include_graphics("figures_tables/C_variable_mapping.png")
```

### 2023-05-23 doc/manuscript/ESSDmanuscript.tex
exclude R_het_soil, knit in new display items
closes #51
closes #45

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned net emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accessible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
inventories if made available through the Emission Factor Database
(EFDB) of the International Panel on Climate Change (IPCC). Here, we
develop and document a process for semi-automated submission of data
from ForC into the EFDB, assess the data available and submitted to
date, and provide recommendations for improving forest data collection,
analysis, and reporting to improve inventories of forest-sector
greenhouse gas emissions and removals. We begin by reconciling
terminology and mapping ForC fields into EFDB. This process required
some updates to the ForC database structure, leading to the release of a
new version of ForC (v4.0; described here). As of May 23, 2023, ForC
contained \textasciitilde19323 independent records relevant to EFDB,
1438 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C inventories.

@@

Aboveground biomass, which is typically \textless200 t C
ha\textsuperscript{-1} but can exceed 700 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by the IPCC as ``all
biomass of living vegetation above the soil including stems, stumps,
branches, bark, seeds, and foliage''
\citep{ipcc_good_2003, ipcc_2006_2006}. IPCC's guidance is that the
understory may be excluded the understory if it constitutes a ``minor''
component, \emph{where quantitative definitions of ``understory'' and
``minor'' are not provided}, but where a commonly applied minimum size
sampling threshold for mature forests would be 10 cm stem diameter at
breast height (DBH). A recent study characterizing the contributions of
trees in different DBH classes to ecosystem C stocks and fluxes found
that trees 1 - 10 cm DBH contributed up to \textasciitilde8\%
aboveground biomass, \textasciitilde17\% aboveground woody net primary
productivity (\(ANPP_{woody.stem}\)), and \textasciitilde20\% woody
mortality (\(M_{woody}\)) of mature closed-canopy forests worldwide
\citep{piponiot_distribution_2022}. In regrowth forests, woodlands, or
savannas, small trees and shrubs contribute a much larger proportion of
C stocks and fluxes
\citep{lutz_global_2018, piponiot_distribution_2022, hughes_biomass_1999},
and, correspondingly, biomass estimates for these ecosystems tend
include smaller size classes. While IPCC guidance specifies that all
living vegetation should be included in biomass estimates, forest
censuses and biomass estimates do not consistently include life forms
other than dicot trees (e.g., lianas, ferns, palms, bamboo), although
these do tend to be censused when they constitute a large proportion of
the biomass \citep[e.g.,][]{fukushima_recovery_2007}. Further, it is
important to note that the IPCC definition of aboveground biomass
excludes standing dead wood, which is included in remote sensing biomass
estimates \citep{duncanson_aboveground_2021}.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C stock change calculations under the Paris Agreement.
However, modest changes to the structure and contents of ForC were
needed in order to provide all information required by EFDB and to
improve ForC's capacity to serve as a repository of valuable information
for forest C inventories under IPCC guidelines. To support export of
data to EFDB, and to improve the overall quality of the ForC database,
we added or modified 18 fields (Appendix A), defined 15 new variables,
implemented enhanced quality control, manually reviewed \textgreater1954
records to obtain additional required information, and added 329 new
records.

@@

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon.}\label{fig:fig_variable_mapping}
\end{figure}

@@

As of May 23, 2023, ForC (v4.0) contained 32693 independent records
(39848 total), 19323 of which were for the 43 variables relevant to EFDB
(Fig. 1). These records were distributed across all forested continents
and ecozones, albeit unevenly (Fig. 2). The largest number of records
came from Asia, followed by North America, South America, and Europe,
with relatively few records from Africa, Australia, and Oceania (Fig.
3c). Categorized by FAO ecozone, the greatest numbers of records came
from subtropical humid forests, temperate mountain systems, and tropical
rain forests, each with \textgreater2,000 independent records (Fig. 3b).
Boreal coniferous forests, temperate continental forests, subtropical
mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 29 of the 43 variables (or closely-related
variable groups) relevant to EFDB (Table 2, Fig. 1). The records were
very unevenly distributed across variables. The variable with most
records was aboveground biomass, representing 42\% of all independent
records relevant to EFDB, and aboveground biomass components (woody
biomass or foliage) representing an additional 5\%. A total of 27\% of
relevant records were for root biomass (including fine and coarse root
components), while 4\% described total biomass. The non-living pools
were less represented, with 4\% of relevant were for dead wood
(including standing and fallen components), 0.4\% for litter, 0.3\% for
total ecosystem C excluding soils, and 2.1\% for soil carbon.

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n sent to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1094 & 850 & 95 & 50\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9456 & 8155 & 1348 & 764\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 520 & 73 & 45\\
\hline
delta.agb & 166 & 150 & 145 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_woody\_stem & 949 & 622 & 60 & 61\\
\hline
ANPP\_woody\_branch & 243 & 200 & 4 & 4\\
\hline
woody.mortality\_ag & 112 & 75 & 47 & 50\\
\hline
stem\_pC & 9 & 0 & 0 & 0\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4185 & 125 & 57\\
\hline
biomass\_root\_fine & 930 & 595 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 413 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root & 577 & 416 & 0 & 0\\
\hline
BNPP\_root\_fine & 488 & 331 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 304 & 104 & 70\\
\hline
deadwood\_standing & 153 & 121 & 18 & 17\\
\hline
deadwood\_down & 425 & 369 & 52 & 28\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 45 & 45 & 45 & 40\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
ANPP\_litterfall & 294 & 253 & 11 & 11\\
\hline
NPP\_litter & 94 & 70 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Total Ecosystem C (excl. soils)} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
total.ecosystem\_2 & 64 & 64 & 0 & 0\\
\hline
delta.total.ecosystem\_2 & 0 & 0 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 401 & 89 & 56\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
R\_het\_soil & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{23575} & \textbf{19323} & \textbf{2283} & \textbf{1438}\\
\hline
\end{longtable}
\endgroup{}

@@

As of May 23, 2023, we had reviewed or added 2283 EFDB-relevant records,
1438 records of which were submitted to EFDB, and 376 of which have been
reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review the records,
and it is not expected that many -- if any -- records will be rejected.

>>>>

\textbf{Abstract.} Forests are critical for climate change mitigation
and constitute a substantial portion of planned net emissions reductions
under the 2015 Paris Agreement. Yet, the efficacy of greenhouse gas
mitigation planning and reporting is dependent upon the quality of
available emission factors data, including forest carbon (C) stocks and
changes therein. Tens of thousands of relevant forest C estimates have
been published, yet are not readily accessible to the practitioners
compiling national greenhouse gas inventories. Many of these data have,
however, been compiled in the Global Forest C database (ForC;
https://forc-db.github.io/) and stand to be of value to greenhouse gas
inventories if made available through the Emission Factor Database
(EFDB) of the International Panel on Climate Change (IPCC). Here, we
develop and document a process for semi-automated submission of data
from ForC into the EFDB, assess the data available and submitted to
date, and provide recommendations for improving forest data collection,
analysis, and reporting to improve inventories of forest-sector
greenhouse gas emissions and removals. We begin by reconciling
terminology and mapping ForC fields into EFDB. This process required
some updates to the ForC database structure, leading to the release of a
new version of ForC (v4.0; described here). As of May 23, 2023, ForC
contained \textasciitilde19316 independent records relevant to EFDB,
1438 of which have been submitted to date. Among the data in ForC, there
is disproportionate representation of biomass (particularly aboveground)
stocks, with far fewer records for dead organic matter and soil C, and
relatively few or no records for net annual increments or C fluxes into
(gains) or out of (losses) the IPCC-defined C pools. Geographic
representation is also quite uneven, with the highest densities of
relevant records in temperate forests, and with relatively scant
representation of tropical forests in Africa and Asia. This distribution
of records is generally reflected in the subset of records that have
been submitted to EFDB to date. In the future, forest C estimates in
EFDB can be improved through targeted research to fill critical gaps,
reporting of information required by IPCC, and continued submission of
data from scientific publications to the EFDB. Given that climate change
is rapidly impacting the world's forests, timely reporting of recent
estimates will be especially critical to accurate forest C inventories.

@@

Aboveground biomass, which is typically \textless200 t C
ha\textsuperscript{-1} but can exceed 700 t C ha\textsuperscript{-1}
\citep{anderson-teixeira_carbon_2021}, is defined by the IPCC as ``all
biomass of living vegetation above the soil including stems, stumps,
branches, bark, seeds, and foliage''
\citep{ipcc_good_2003, ipcc_2006_2006}. IPCC's guidance is that the
understory may be excluded the understory if it constitutes a ``minor''
component (defined as \ldots{}), and where a commonly applied minimum
size sampling threshold for mature forests would be 10 cm stem diameter
at breast height (DBH). A recent study characterizing the contributions
of trees in different DBH classes to ecosystem C stocks and fluxes found
that trees 1 - 10 cm DBH contributed up to \textasciitilde8\%
aboveground biomass, \textasciitilde17\% aboveground woody net primary
productivity (\(ANPP_{woody.stem}\)), and \textasciitilde20\% woody
mortality (\(M_{woody}\)) of mature closed-canopy forests worldwide
\citep{piponiot_distribution_2022}. In regrowth forests, woodlands, or
savannas, small trees and shrubs contribute a much larger proportion of
C stocks and fluxes
\citep{lutz_global_2018, piponiot_distribution_2022, hughes_biomass_1999},
and, correspondingly, biomass estimates for these ecosystems tend
include smaller size classes. While IPCC guidance specifies that all
living vegetation should be included in biomass estimates, forest
censuses and biomass estimates do not consistently include life forms
other than dicot trees (e.g., lianas, ferns, palms, bamboo), although
these do tend to be censused when they constitute a large proportion of
the biomass \citep[e.g.,][]{fukushima_recovery_2007}. Further, it is
important to note that the IPCC definition of aboveground biomass
excludes standing dead wood, which is included in remote sensing biomass
estimates \citep{duncanson_aboveground_2021}.

@@

Previous versions of ForC
\citep{anderson-teixeira_carbon_2016, anderson-teixeira_forc_2018, anderson-teixeira_carbon_2021}
contained most of the information required by EFDB, and, more broadly,
to inform C stock change calculations under the Paris Agreement.
However, modest changes to the structure and contents of ForC were
needed in order to provide all information required by EFDB and to
improve ForC's capacity to serve as a repository of valuable information
for forest C inventories under IPCC guidelines. To support export of
data to EFDB, and to improve the overall quality of the ForC database,
we added or modified 18 fields (Appendix A), defined 15 new variables,
implemented enhanced quality control, manually reviewed \textgreater1963
records to obtain additional required information, and added 329 new
records.

@@

\begin{figure}
\includegraphics[width=14cm]{figures_tables/C_variable_mapping} \caption{\textbf{Schematic illustrating the carbon pools defined under IPCC Guidelines for national greenhouse gas inventories; corresponding ForC variables, and relationships among them.} For each C pool, we show ForC variables corresponding to the stock, stock change (net annual increment), gain (influx), and loss. Most, but not all, EFDB-relevant ForC variables are shown here. Correspondence of ForC variables to IPCC criteria often depends upon measurement protocols (e.g., minimum stem diameter censused). Additional caveats are as follows: (a,b) branch fall and mortality of stems below the minimum stem diameter censused, which are necessary for a full accounting of dead organic matter production but typically assumed negligible for calculations of biomass change, are excluded by common measurement practice (a) or ForC variable definition (b); (c) assumes that leaf production equals leaf fall, or that changes in foliage biomass are negligble; (d,e) belowground components excluded by common measurement practice (d) or ForC variable definition (e); (f) excludes movement of dead wood into litter through breakage or size reduction; (g) measurements often limited to litter horizon (OL) and may exclude larger branches and stems classified as litter and/or the more decomposed layers of the O horizon. **This variable is techically EFDB-relevant but not selected for submission because their is no corresponding influx variable.}\label{fig:fig_variable_mapping}
\end{figure}

@@

As of May 23, 2023, ForC (v4.0) contained 32693 independent records
(39848 total), 19316 of which were for the 42 variables relevant to EFDB
(Fig. 1). These records were distributed across all forested continents
and ecozones, albeit unevenly (Fig. 2). The largest number of records
came from Asia, followed by North America, South America, and Europe,
with relatively few records from Africa, Australia, and Oceania (Fig.
3c). Categorized by FAO ecozone, the greatest numbers of records came
from subtropical humid forests, temperate mountain systems, and tropical
rain forests, each with \textgreater2,000 independent records (Fig. 3b).
Boreal coniferous forests, temperate continental forests, subtropical
mountain systems, and tropical moist deciduous forests had
\textgreater1,000 independent records each, while other ecozones all had
\textless1,000 records. The most widely represented forest type was
needleleaf evergreen, followed by broadleaf deciduous and broadleaf
evergreen (Fig. 3a). In terms of stand age, the most represented age
class was 20-100 years, followed by \textless20 years and then
\textgreater100 years (Fig. 3d).

@@

ForC contained records for 29 of the 42 variables (or closely-related
variable groups) relevant to EFDB (Table 2, Fig. 1). The records were
very unevenly distributed across variables. The variable with most
records was aboveground biomass, representing 42\% of all independent
records relevant to EFDB, and aboveground biomass components (woody
biomass or foliage) representing an additional 5\%. A total of 27\% of
relevant records were for root biomass (including fine and coarse root
components), while 4\% described total biomass. The non-living pools
were less represented, with 4\% of relevant were for dead wood
(including standing and fallen components), 0.4\% for litter, 0.3\% for
total ecosystem C excluding soils, and 2.1\% for soil carbon.

@@

\begin{longtable}[t]{l|l|l|l|l}
\caption{\label{tab:table_variables}\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n submitted to EFDB}\\
\hline
\endfirsthead
\caption[]{\textbf{Numbers of records of ForC variables (or closely related variable groups) relevant to, and sent to, EFDB.} \textit{(continued)}}\\
\hline
\textbf{variable} & \textbf{n in ForC} & \textbf{n independent records in ForC} & \textbf{n reviewed} & \textbf{n submitted to EFDB}\\
\hline
\endhead
\textbf{Biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass & 1094 & 850 & 95 & 50\\
\hline
delta.biomass & 0 & 0 & 0 & 0\\
\hline
NPP\_woody & 136 & 93 & 0 & 0\\
\hline
woody.mortality & 0 & 0 & 0 & 0\\
\hline
\textbf{Aboveground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_ag & 9449 & 8148 & 1357 & 764\\
\hline
biomass\_ag\_woody & 460 & 366 & 10 & 10\\
\hline
biomass\_ag\_foliage & 601 & 520 & 73 & 45\\
\hline
delta.agb & 166 & 150 & 145 & 123\\
\hline
ANPP\_woody & 299 & 242 & 0 & 0\\
\hline
ANPP\_woody\_stem & 949 & 622 & 60 & 61\\
\hline
ANPP\_woody\_branch & 243 & 200 & 4 & 4\\
\hline
woody.mortality\_ag & 112 & 75 & 47 & 50\\
\hline
stem\_pC & 9 & 0 & 0 & 0\\
\hline
\textbf{Belowground biomass} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
biomass\_root & 4629 & 4185 & 125 & 57\\
\hline
biomass\_root\_fine & 930 & 595 & 18 & 18\\
\hline
biomass\_root\_coarse & 599 & 413 & 12 & 7\\
\hline
delta.biomass\_root & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_coarse & 0 & 0 & 0 & 0\\
\hline
delta.biomass\_root\_fine & 0 & 0 & 0 & 0\\
\hline
woody.mortality\_root & 0 & 0 & 0 & 0\\
\hline
BNPP\_root & 577 & 416 & 0 & 0\\
\hline
BNPP\_root\_fine & 488 & 331 & 0 & 0\\
\hline
BNPP\_root.turnover\_fine & 91 & 56 & 0 & 0\\
\hline
BNPP\_root\_coarse & 329 & 250 & 0 & 0\\
\hline
\textbf{Dead wood} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
deadwood & 438 & 304 & 104 & 70\\
\hline
deadwood\_standing & 153 & 121 & 18 & 17\\
\hline
deadwood\_down & 425 & 369 & 52 & 28\\
\hline
delta.deadwood & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_standing & 0 & 0 & 0 & 0\\
\hline
delta.deadwood\_down & 0 & 0 & 0 & 0\\
\hline
R\_het\_deadwood & 0 & 0 & 0 & 0\\
\hline
\textbf{Litter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
O.horizon & 45 & 45 & 45 & 40\\
\hline
delta.O.horizon & 4 & 4 & 4 & 4\\
\hline
litter & 30 & 30 & 23 & 23\\
\hline
delta.litter & 0 & 0 & 0 & 0\\
\hline
ANPP\_litterfall & 294 & 253 & 11 & 11\\
\hline
NPP\_litter & 94 & 70 & 0 & 0\\
\hline
R\_het\_litter & 167 & 143 & 0 & 0\\
\hline
\textbf{Total Ecosystem C (excl. soils)} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
total.ecosystem\_2 & 64 & 64 & 0 & 0\\
\hline
delta.total.ecosystem\_2 & 0 & 0 & 0 & 0\\
\hline
\textbf{Soil organic matter} & \textbf{} & \textbf{} & \textbf{} & \textbf{}\\
\hline
SOM / SOC & 693 & 401 & 89 & 56\\
\hline
delta.SOM / delta.SOC & 0 & 0 & 0 & 0\\
\hline
\textbf{TOTAL} & \textbf{23568} & \textbf{19316} & \textbf{2292} & \textbf{1438}\\
\hline
\end{longtable}
\endgroup{}

@@

As of May 23, 2023, we had reviewed or added 2292 EFDB-relevant records,
1438 records of which were submitted to EFDB, and 376 of which have been
reviewed, accepted, and posted (Figs. 2-3, Table 2). The 37\%
attenuation between records reviewed and those sent to EFDB was
attributable to the presence of digitized records and records where a
variable's value had been calculated as the sum or difference of related
variables rather than presented directly in the text. The discrepancy
between the number of records sent and that posted to EFDB is primarily
attributable to the time required for the IPCC to review the records,
and it is not expected that many -- if any -- records will be rejected.
